---
title: "Deploying aurora apps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Deploying aurora apps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Because an aurora app is stateless, deployment is the boring, scalable kind: a
container that serves static assets plus JSON routes. Run as many replicas as
you like behind a load balancer — there are no sticky sessions to worry about.

## Build an image

The static UI is compiled at build time and shipped as `www/index.html`; the
container **serves** it and does not rebuild it, so the runtime image installs
no UI dependencies (bslib, and transitively shiny). Build the UI before the
image:

```{r}
library(aurora)

aurora_build_ui("meu_app")   # compile www/index.html (needs bslib; run on dev/CI)

# Generate a Dockerfile (+ .dockerignore). It installs only the runtime deps
# your routers/helpers need, plus plumber2 and aurora.
aurora_dockerfile("meu_app")

# Build (and optionally push) the image using the docker CLI.
aurora_build_image("meu_app", tag = "org/meu_app:latest", push = TRUE)
```

The generated Dockerfile pulls R packages as prebuilt **binaries** from Posit
Package Manager, so builds are fast and need no compiler toolchain at run time.
Posit Package Manager only serves **amd64 (x86_64)** Linux binaries, so on an
Apple Silicon machine build for the production target explicitly — otherwise the
native arm64 build compiles every dependency from source:

```sh
docker build --platform linux/amd64 -t org/meu_app:latest .
```

`aurora_dockerfile()` writes a Dockerfile whose entry point is `Rscript api.R`,
so the container and local development share one assembly path. Key arguments:

- `flavor` — `"debian"` (default) or `"alpine"` (see below).
- `base` — base image; `NULL` resolves per flavor (`rocker/r-ver` / `rhub/r-minimal`).
- `sysdeps` — `"auto"` uses a curated default set covering the plumber2 + bslib
  baseline plus common TLS/curl/db/geo/graphics needs; pass a vector to override.
- `port` — exposed port (default `8000`).

### Choosing a flavor

| | `debian` (default) | `alpine` |
|---|---|---|
| Base | `rocker/r-ver` | `rhub/r-minimal` |
| R packages | **binaries** from Posit Package Manager (fast) | **compiled** from source via `installr` (slower) |
| Image size | larger | tiny (~25 MB base) |
| Arch | amd64 binaries (arm64 compiles) | builds natively on amd64 **and** arm64 |
| Best for | heavy/geo apps, fast CI, broad compatibility | size-sensitive / edge deploys, simple deps |

```{r}
aurora_dockerfile("meu_app", flavor = "alpine")
```

The `alpine` flavor compiles everything (no CRAN binaries on Alpine) and uses
`installr -d -t "<build deps>" -a "<runtime libs>"`. aurora ships defaults that
cover the plumber2 + bslib baseline; for extra system libraries (e.g. GDAL/GEOS
for `sf`) pass them via `sysdeps`. Note that aurora's baseline still pulls a
non-trivial tree (httpuv, the fiery stack, roxygen2, the graphics packages), so
even a small app compiles a fair amount on Alpine — the win is final image size.

## Runtime configuration (environment variables)

The generated `api.R` reads its bind address and port from the environment, and
aurora features are env-toggleable, so the **same image** runs in dev and prod:

| Variable | Used for |
|---|---|
| `AURORA_HOST` / `AURORA_PORT` | bind address / port (`api.R`) |
| `AURORA_OTEL` | enable OpenTelemetry logging (`vignette("telemetry")`) |
| `AURORA_JWT_SECRET` | signing secret for the `auth` template |
| `AURORA_ENV=prod` | `Secure; SameSite=Strict` auth cookies (behind HTTPS) |

Never bake secrets into the image — inject them at run time:

```sh
docker run -p 8000:8000 \
  -e AURORA_JWT_SECRET="$(openssl rand -hex 32)" \
  -e AURORA_ENV=prod \
  org/meu_app:latest
```

## Behind a reverse proxy / load balancer

Serve the app under a path prefix or a subdomain via your proxy (nginx, Traefik,
an ingress). The runtime resolves API paths against the page's base path, so an
app served under `/meu_app/` still calls its routes correctly. Run multiple
replicas freely — state lives in the client (cookies) or an external store, not
in the R process (see `vignette("aurora")` on `aurora_data_store()`).

## ShinyProxy

ShinyProxy launches the container like any Docker-backed app.
`aurora_shinyproxy_yaml()` emits the `proxy.specs` entry for you:

```{r}
aurora_shinyproxy_yaml(
  image = "org/meu_app:latest",
  dir   = "meu_app",              # defaults id / display-name from the app name
  env   = list(AURORA_ENV = "prod")
)
#> - id: meu_app
#>   display-name: meu_app
#>   container-image: org/meu_app:latest
#>   port: 8000
#>   container-env:
#>     AURORA_ENV: prod
```

Paste that under `proxy.specs` in your ShinyProxy config, or pass `wrap = TRUE`
for a complete `proxy: specs:` snippet (and `write = TRUE` to save it to a file).

## Checklist

- [ ] Strong `AURORA_JWT_SECRET` injected at run time (if using auth).
- [ ] `AURORA_ENV=prod` and HTTPS terminated at the proxy.
- [ ] Image rebuilt after dependency changes (sysdeps are resolved at build).
- [ ] Health check wired to a public route (e.g. `/health`).
